---
title: "How to use opencis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use opencis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# Necessary in order to pass the re-building of the vignette step of the R-CMD-check action on Ubuntu systems
if(Sys.info()["sysname"] == "Linux") {
  httr::set_config(httr::config(ssl_verifypeer = 0L)) 
}
```

## Installation

You can install `opencis` from Github using the `devtools` package:

``` {r, eval=FALSE}
devtools::install_github("hmeleiro/opencis")
```


## Usage

The package includes three group of functions. 

The first group is used to search for surveys, survey questions and timeseries. 

``` {r}
library(opencis)

# Search for survey studies
search_studies(search_terms = "preelectoral", since_date = "01-01-2020", until_date = "17-11-2023")

# Search for survey questions
search_questions(search_terms = "feminismo")

# Search for timeseries
search_series(search_terms = "situación económica")
```

The second group of functions retrieves the valid category codes to narrow search queries. [Read the documentation](https://hmeleiro.github.io/opencis/reference/index.html#retrieve-categories) to understand how to use the category codes in the search functions.

``` {r}
get_study_categories()

get_question_categories()

get_series_category()

```


The third group of functions are used to import the data directly into R. 

```{undefined warning=FALSE}
# Read a survey study
df <- read_cis("3411")
print(df)


# Read a timeseries
df_series <- read_series("2092")
print(df_series)
```
