---
title: "How to use opencis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{How to use opencis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

## Installation

You can install `opencis` from Github using the `devtools` package:

```{r, eval=FALSE}
devtools::install_github("hmeleiro/opencis")
```

## Usage

### Searching for studies, questions and series

`search_cis()` searches the CIS catalogue and returns a tibble with matching results. The `catalogo` argument controls what type of item is searched: `"estudio"` (default), `"pregunta"` or `"serie"`. You can restrict results to a date range with `from` and `to`, and change the sort order with `sort` (`"relevance"`, `"publishDate-"`, `"publishDate+"`).

```{r, eval = FALSE}
library(opencis)

# Search for survey studies
search_cis(q = "preelectoral", from = "2020-01-01", to = "2023-11-17")

# Search for survey questions
search_cis(q = "feminismo", catalogo = "pregunta")

# Search for data series
search_cis(q = "situación económica", catalogo = "serie")
```

By default `search_cis()` returns only the first page of results. Use `search_all_cis()` to automatically paginate through all pages and get every matching result in a single tibble:

```{r, eval = FALSE}
# Retrieve all postelectoral studies (all pages)
all_studies <- search_all_cis(q = "postelectoral")
print(nrow(all_studies))

# Filter by date range across all pages
studies <- search_all_cis(
  q    = "ideologia",
  from = "2010-01-01",
  to   = "2020-12-31"
)
```

`search_all_cis()` accepts the same arguments as `search_cis()`.

------------------------------------------------------------------------

### Reading study data into R

`read_cis()` downloads the SPSS data file for a study and imports it directly into R as a labelled data frame (via `haven`):

```{r, eval = FALSE}
df <- read_cis(3411)
print(df)
```

------------------------------------------------------------------------

### Exploring variables: the data dictionary

After loading a study with `read_cis()`, use `get_data_dictionary()` to obtain a tidy tibble with every variable name, its label and its value labels:

```{r, eval = FALSE}
df   <- read_cis(3328)
dict <- get_data_dictionary(df)
print(dict)

# Find variables whose label contains a keyword
dict[grepl("sexo", dict$label, ignore.case = TRUE), ]

# Inspect value labels for a specific variable
dict$value_labels[[which(dict$variable == "SEXO")]]
```

------------------------------------------------------------------------

### Getting study metadata

`get_metadata()` retrieves the technical information sheet of a study from the CIS website — field dates, study type, country, authorship, thematic indices, etc. — and returns it as a two-column tibble (`field`, `value`):

```{r, eval = FALSE}
meta <- get_metadata(3328)
print(meta)

```

------------------------------------------------------------------------

### Downloading the ZIP file to disk

If you want to keep the raw data files instead of reading them into a temporary directory, use `download_study()`. It saves the ZIP archive to any local folder:

```{r, eval = FALSE}
# Save to the current working directory
path <- download_study(3328)
cat("Saved to:", path, "\n")

# Save to a specific folder
path <- download_study(3328, destdir = "data/raw")
cat("Saved to:", path, "\n")
```

------------------------------------------------------------------------

### Browsing the questionnaire and technical sheet

`browse_pdf()` extracts the PDF documents bundled inside the study ZIP and opens them in your default browser. CIS ZIPs typically include two PDFs:

- **Questionnaire** (`wanted_file = "cues"`, default)
- **Technical sheet** (`wanted_file = "ft"`)

```{r, eval = FALSE}
# Open the questionnaire for study 3328
browse_pdf(3328)

# Open the technical sheet
browse_pdf(3328, wanted_file = "ft")
```
